Reproducible Research Project 1
=========================================

# Author: Michael Cui

## 1. Loading the dataset
```{r loaddataset}
setwd("~/Downloads")
data <- read.csv("activity.csv")
```

## 2. Histogram of the total number of steps taken each day
```{r totalsteps}
library(ggplot2)
dailysum <- tapply(data$steps, data$date, sum, na.rm = TRUE)
qplot(dailysum, xlab = "total number of steps taken each day")
```

## 3. Mean and median number of steps taken each day
```{r summary}
summary(dailysum)
```

## 4. Time series plot of the average number of steps taken
```{r stepsbytime}
library(ggplot2)
averagesteps <- aggregate(x=list(steps=data$steps), by=list(interval=data$interval)                               ,FUN=mean, na.rm=TRUE)
ggplot(data = averagesteps, aes(x = interval, y = steps)) + 
                            geom_line() +
                            ylab("steps") +
                            xlab("interval")
```

## 5. The 5-minute interval that, on average, contains the maximum number of steps
```{r maxinterval}
maxvalue <- which.max(averagesteps$steps)
averagesteps[maxvalue, ]
```

## 6. Code to describe and show a strategy for imputing missing data
```{r nadescription}
nas <- is.na(data$steps)
sum(nas) ## total count of NAs
mean(nas) ## percentage of NAs
```
We can see that there is a significant amount of NAs. This might influence the results. To mitigate this effect, we need to input missing data. 
First, find out where the missing values are. 
```{r nasummary}
missing <- data[nas, ]
## observe the data
summary(missing)
unique(missing$date)
```
We can see that missing values are located in 8 different days, each with 288 missing values. There is a total of 2304 missing values.
Now we fill each NA with the mean number of steps in that specific interval.
```{r fillna}
library(dplyr)
averagesteps2 <- as.list(averagesteps$steps)
missing$steps <- rep(averagesteps2, 8)
fulldata <- data[!nas, ]
inputeddata <- rbind(fulldata, missing)
inputeddata <- arrange(inputeddata, date)
inputeddata <- as.data.frame(inputeddata)
```
## 7. Histogram of the total number of steps taken each day after missing values are imputed
```{r totalsteps2}
library(ggplot2)
inputeddata$steps <- as.numeric(inputeddata$steps)
dailysum2 <- tapply(inputeddata$steps, inputeddata$date, sum, na.rm = TRUE)
qplot(dailysum2, xlab = "total number of steps taken each day")
```
## 8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r weekdayplot}
library(lubridate)
inputeddata$date <- as.POSIXct(inputeddata$date)
inputeddata_weekday <- mutate(inputeddata, weekday=weekdays(date))
dailysum2_weekday_interval <- group_by(inputeddata_weekday, weekday, interval)
results <- summarise(dailysum2_weekday_interval, steps=mean(steps, na.rm=TRUE))
qplot(interval, steps, data=results, facets= weekday ~ .) + geom_line()
```
